\section{Durchführung}

\subsection{Maturity}
\begin{itemize}
    \item Technologiereifegrad: technology readyness level
    \item bekannte Fehler bzw.\ ungelöste Probleme
    \item Stärke der Dokumentation
    \item Anzahl der aktuellen Nutzer (github stars/forks oder so)
    \item werden moderne Sprachfeatures genutzt
\end{itemize}

\paragraph{Nova}
first stable release: 2018-08-22

major versions: 4

\paragraph{filament}
first stable release: 2021-03-02

major versions: 2

\subsubsection{Roadmap}
wie aktiv wird entwickelt

nova: no roadmap

filament: yes, need to check

\subsection{Kostenvergleich}
tbd

\subsection{Community}
> https://www.reddit.com/r/laravel/comments/wn8n8u/laravel_nova_vs_filament/

- I have worked with both and found Filament much easier to work with.
- Another problem I encountered with Nova is that customers without a technical background have a difficult time to work
with Nova.
- Customising nova is really difficult.
- I've used both, and i still use both , but i prefer filament for it's community and easy customization, but i still
use nova for vue projects.
- Nova is perfect if you don't need customization
- In my experience if you don’t need alot of custom work use Nova cause it’s easier and way faster to develop with.
- If you need something custom use filament.
- Filament is a lot easier to customize and extend. Mainly because you don’t have the compile Vue, etc.

> https://world.hey.com/akviby/filament-is-the-best-thing-that-happened-to-me-since-laravel-nova-2ab28a0f

> https://laracasts.com/discuss/channels/laravel/whats-the-best-free-laravel-admin-panel

\newpage

\subsection{Vergleich auf Basis der Features}
\input{sections/45_feature_comparison}

\subsection{Prototypischer Vergleich}
konkret: können die Probleme mit der Flight View optimiert werden?

\subsubsection{Positive Erfahrungen}
tbd

\subsubsection{Probleme}
tbd

\subsection{Code Style}
\begin{itemize}
    \item PSR Compliance
    \item clean code literatur nach robert martin
    \item wie streng getyped sind die frameworks
\end{itemize}

\subsection{Vergleich anhand von Design Principles}
tbd

\subsubsection{SOLID}
tbd

\subsubsection{KISS Principle}
tbd

\subsubsection{DRY Principle}
tbd

\subsection{Vergleich anhand von Design Patterns}
\begin{itemize}
    \item hier ist ein Problem: wurde das mit Muster gelöst
    \item sind Muster dokumentiert
    \item ein Muster oder viele Muster
    \item sind Muster angemessen eingesetzt
    \item Qualität wichtiger Quantität
    \item muster quantitativ bewerten
    \item bei einem muster mal qualitativ in die tiefe gehen
\end{itemize}
